Performance-portable Parallel Programming in Computer Science Education with Kokkos

### 2020-01-23 src/2.GenericProgramming.tex
Added Hollman's pic.
More typos

Lastly, a model must decide between declarative and imperative semantics. Declarative semantics require a developer to express intent, implementation is the responsibility of the programming model and underlying toolchain. Imperative semantics require the developer to specify the mechanisms by which their intent will be implemented. In an imperative model, developer success is tied to their knowledge of implementation mechanisms. In a declarative model that success is tied to the developer\'s ability to accurately declare their requirements, and the toolchain\'s ability to turn that into an implementation. Figures~\ref{figOMPLike} and~\ref{figKokkosLike} show examples of two sample applications using the aforementioned types of semantics. While both programming models implement the same programming paradigm (parallel patterns), Figure~\ref{figOMPLike} required the developer to express how to parallelize the loop, while in~\ref{figKokkosLike} the developer just declared their parallel semantics, and relied on the programming model to pick an implementation.

Kokkos takes the declarative approach. For the developers of our large scientific applications, this is appealing because it allows them to express their algorithms in one way, and rely on Kokkos to map that intent effectively to a variety of architectures. For students this is appealing because they\'re unlikely to have expertise in these architectures. For an educator looking for an implementation vehicle with which to teach students about parallelism, the declarative model eliminates the burden of reimplementing exercises as computer labs and dominant architectures change. For a declarative model to make sense, the educator should know something about the abstract execution model being used to ensure performance portability. Depending on the nature of the course, this could be mostly hidden from students, or discussed in detail.

>>>>

Lastly, a model must decide between declarative and imperative semantics. Declarative semantics require a developer to express intent, implementation is the responsibility of the programming model and underlying toolchain. Imperative semantics require the developer to specify the mechanisms by which their intent will be implemented. In an imperative model, developer success is tied to their knowledge of implementation mechanisms. In a declarative model that success is tied to the developer's ability to accurately declare their requirements, and the toolchain's ability to turn that into an implementation. Figures~\ref{figOMPLike} and~\ref{figKokkosLike} show examples of two sample applications using the aforementioned types of semantics. While both programming models implement the same programming paradigm (parallel patterns), Figure~\ref{figOMPLike} required the developer to express how to parallelize the loop, while in~\ref{figKokkosLike} the developer just declared their parallel semantics, and relied on the programming model to pick an implementation.

Kokkos takes the declarative approach. For the developers of our large scientific applications, this is appealing because it allows them to express their algorithms in one way, and rely on Kokkos to map that intent effectively to a variety of architectures. For students this is appealing because they're unlikely to have expertise in these architectures. For an educator looking for an implementation vehicle with which to teach students about parallelism, the declarative model eliminates the burden of reimplementing exercises as computer labs and dominant architectures change. For a declarative model to make sense, the educator should know something about the abstract execution model being used to ensure performance portability. Depending on the nature of the course, this could be mostly hidden from students, or discussed in detail.

### 2020-01-23 src/8.Related.tex
More typos

\section{Related Work}\label{chap:related}
\subsection{RAJA}
Like Kokkos, RAJA is a C++ framework for porting massive DOE codes to a variety of architectures. In both frameworks, a developer first expresses the operations their kernel performs using simple C++ functional programming mechanisms. The difference is that Kokkos is declarative, while RAJA is imperative. In Kokkos, the framework figures out how to map that kernel to different architectures. In RAJA, functionality is provided to make it easy for a developer to try different strategies to map those kernels to architectures, but the developer still decides how that mapping happens. Which model makes more sense in a given circumstance depends on how much exposure to the underlying programming model a developer wants. RAJA would be ideal for a class in which the educator wants to expose students to the semantics of different programming models without exposing them to the detailed syntax of those models.
\subsection{DPC++}
DPC++ is a model developed by Intel(TM) for expressing parallelism across their architectures. In addition to sharing a declarative philosophy with Kokkos, it even share a number of patterns. Unlike Kokkos, at time of publication it targets only Intel architectures (it isn't vendor-agnostic) For a course targeting Intel(TM) platforms, you\'re likely to see good tool support, DPC++ could be a valid option.
\subsection{Parallel Constructs in the C++ Standard}
C++ increasingly supports parallel constructs. We discuss the contributions of Kokkos to C++ in~\ref{chap:c++}, it's increasingly possible to teach parallelism with nothing but a C++ compiler. A problem is that at time of publication C++ doesn't support accelerators, or many of the complexities of parallel computing as it\'s used in practice. A positive is that instead of requiring libraries, you just need a modern C++ compiler. For a course in which topics like accelerators aren\'t discussed, this can be a valid option.
%\end{itemize}

>>>>

\section{Related Work}\label{chap:related}
\subsection{RAJA}
Like Kokkos, RAJA is a C++ framework for porting massive DOE codes to a variety of architectures. In both frameworks, a developer first expresses the operations their kernel performs using simple C++ functional programming mechanisms. The difference is that Kokkos is declarative, while RAJA is imperative. In Kokkos, the framework figures out how to map that kernel to different architectures. In RAJA, functionality is provided to make it easy for a developer to try different strategies to map those kernels to architectures, but the developer still decides how that mapping happens. Which model makes more sense in a given circumstance depends on how much exposure to the underlying programming model a developer wants. RAJA would be ideal for a class in which the educator wants to expose students to the semantics of different programming models without exposing them to the detailed syntax of those models.
\subsection{DPC++}
DPC++ is a model developed by Intel(TM) for expressing parallelism across their architectures. In addition to sharing a declarative philosophy with Kokkos, it even share a number of patterns. Unlike Kokkos, at time of publication it targets only Intel architectures (it isn't vendor-agnostic) For a course targeting Intel(TM) platforms, you're likely to see good tool support, DPC++ could be a valid option.
\subsection{Parallel Constructs in the C++ Standard}
C++ increasingly supports parallel constructs. We discuss the contributions of Kokkos to C++ in~\ref{chap:c++}, it's increasingly possible to teach parallelism with nothing but a C++ compiler. A problem is that at time of publication C++ doesn't support accelerators, or many of the complexities of parallel computing as it's used in practice. A positive is that instead of requiring libraries, you just need a modern C++ compiler. For a course in which topics like accelerators aren't discussed, this can be a valid option.
%\end{itemize}
